---
title: "Race Analysis Report - `r format(Sys.time(), '%Y-%m-%d')`"
author: "SailingRaceTrackers"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M')`"
---

```{r setup, include=FALSE, echo=FALSE}
if (!require("gt")) install.packages("gt")
if (!require("dplyr")) install.packages("dplyr")
library(gt)
library(dplyr)
library(jsonlite)
library(lubridate)

COMPUTATION_TIMES_FILE <- "../data/computation_times.txt"
TARGET_BOAT <- "TUF TUF TUF Coeur en liberté"

get_computation_time <- function() {
    if (file.exists(COMPUTATION_TIMES_FILE)) {
        times <- readLines(COMPUTATION_TIMES_FILE)
        if (length(times) > 0) {
            return(ymd_hms(tail(times, 1)))
        }
    }
    return(NULL)
}

save_computation_time <- function() {
    current_time <- format(Sys.time(), "%Y-%m-%d %H:%M:%S")
    write(current_time, file = COMPUTATION_TIMES_FILE, append = TRUE)
}

parse_history_entry <- function(entry) {
    lines_list <- entry$lines[[1]]
    if (is.null(lines_list) || length(lines_list) == 0) return(NULL)
    
    data.frame(
        boat = sapply(lines_list, function(x) as.integer(x[[1]])),
        rank = sapply(lines_list, function(x) as.integer(x[[3]])),
        speed = sapply(lines_list, function(x) as.numeric(x[[9]])),
        vmg = sapply(lines_list, function(x) as.numeric(x[[10]])),
        dtf = sapply(lines_list, function(x) as.numeric(x[[5]])),
        dtl = sapply(lines_list, function(x) as.numeric(x[[6]])),
        dist4h = sapply(lines_list, function(x) as.numeric(x[[12]])),
        dist24h = sapply(lines_list, function(x) as.numeric(x[[16]])),
        heading = sapply(lines_list, function(x) as.numeric(x[[8]])),
        stringsAsFactors = FALSE
    )
}

get_history_for_period <- function(history_list, target_time_str, max_hours = NULL) {
    target_dt <- ymd_hms(target_time_str)
    
    valid_entries <- list()
    for (i in seq_along(history_list)) {
        entry <- history_list[[i]]
        entry_date <- ymd_hms(entry$date)
        hours_diff <- as.numeric(difftime(target_dt, entry_date, units = "hours"))
        
        if (!is.null(max_hours)) {
            if (hours_diff >= 0 && hours_diff <= max_hours) {
                valid_entries[[length(valid_entries) + 1]] <- list(data = entry, hours = hours_diff)
            }
        } else {
            if (hours_diff >= 0) {
                valid_entries[[length(valid_entries) + 1]] <- list(data = entry, hours = hours_diff)
            }
        }
    }
    
    if (length(valid_entries) == 0) return(NULL)
    
    best <- valid_entries[[which.min(sapply(valid_entries, function(x) x$hours))]]
    list(data = parse_history_entry(best$data), hours_ago = best$hours)
}
```

```{r load-data, echo=FALSE}
results <- fromJSON(txt = readLines("../data/boats.json", warn = FALSE))
binfo <- fromJSON("../data/boatinfo.json")

history_df <- results$reports$history
n_history <- nrow(history_df)
latest <- history_df[n_history, ]
latest_time <- latest$date
current_data <- parse_history_entry(latest)

history_list <- list()
for (i in 1:n_history) {
    history_list[[i]] <- list(
        id = history_df$id[i],
        date = history_df$date[i],
        offset = history_df$offset[i],
        lines = history_df$lines[i]
    )
}

binfo_df <- data.frame(
    boat = as.integer(names(binfo)),
    boatName = sapply(binfo, function(x) x$boatName),
    category = sapply(binfo, function(x) x$category),
    boatClass = sapply(binfo, function(x) x$boatClass),
    skipperNames = sapply(binfo, function(x) x$skipperNames)
)

results_df <- left_join(current_data, binfo_df, by = "boat")
df_sorted <- results_df[order(results_df$rank), ]

last_comp_time <- get_computation_time()

data_since_last <- get_history_for_period(history_list, latest_time, max_hours = 0.5)
data_4h <- get_history_for_period(history_list, latest_time, max_hours = 4)
data_24h <- get_history_for_period(history_list, latest_time, max_hours = 24)

save_computation_time()
```

# Current Status and Ranking

**Report generated:** `r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`  
**Data timestamp:** `r latest_time`  
**Last computation:** `r if(!is.null(last_comp_time)) format(last_comp_time, '%Y-%m-%d %H:%M:%S') else "First run"`

## TUF TUF Position

```{r tuf-position, echo=FALSE, results='asis'}
tuf_row <- subset(results_df, grepl("TUF TUF", boatName, ignore.case = TRUE))

if(nrow(tuf_row) > 0 && "boatName" %in% names(tuf_row)) {
    tuf_boat_id <- tuf_row$boat
    tuf_result <- fromJSON(txt = readLines("../data/boats_result.json", warn = FALSE))
    tuf_data <- tuf_result$result[[as.character(tuf_boat_id)]]
    tuf_track <- tuf_data$track
    n_rows <- nrow(tuf_track)
    last_lat <- as.numeric(tuf_track[n_rows, 1])
    last_lon <- as.numeric(tuf_track[n_rows, 2])
    ts <- as.POSIXct(tuf_data$timestamp, origin = "1970-01-01", tz = "UTC")
    
    lat_d <- floor(abs(last_lat))
    lat_m <- round((abs(last_lat) - lat_d) * 60, 1)
    lat_dir <- ifelse(last_lat >= 0, "N", "S")
    lon_d <- floor(abs(last_lon))
    lon_m <- round((abs(last_lon) - lon_d) * 60, 1)
    lon_dir <- ifelse(last_lon >= 0, "E", "W")
    
    cat(paste0("**GPS:** ", lat_d, "° ", lat_m, "' ", lat_dir, "  ", lon_d, "° ", lon_m, "' ", lon_dir, "  \n"))
    cat(paste0("**Position time:** ", format(ts, "%Y-%m-%d %H:%M"), " UTC  \n"))
    cat(paste0("**", tuf_row$boatName, "** - Rank: **", tuf_row$rank, "** / ", nrow(results_df), "  \n"))
    cat(paste0("Speed: ", tuf_row$speed, " kts | VMG: ", tuf_row$vmg, " kts  \n"))
    cat(paste0("DTF: ", tuf_row$dtf, " nm | DTL: ", tuf_row$dtl, " nm  \n"))
    
    if (!is.null(data_4h) && !is.null(data_4h$data)) {
        tuf_4h <- subset(data_4h$data, boat == tuf_row$boat)
        if (nrow(tuf_4h) > 0) {
            cat(paste0("  \n4h ago - Rank: ", tuf_4h$rank, " | Rank change: ", tuf_row$rank - tuf_4h$rank, "  \n"))
        }
    }
    
    if (!is.null(data_24h) && !is.null(data_24h$data)) {
        tuf_24h <- subset(data_24h$data, boat == tuf_row$boat)
        if (nrow(tuf_24h) > 0) {
            cat(paste0("24h ago - Rank: ", tuf_24h$rank, " | Rank change: ", tuf_row$rank - tuf_24h$rank, "  \n"))
        }
    }
}
```

## Overall Leaderboard

```{r table1, echo=FALSE}
df_display <- df_sorted %>%
    mutate(
        boat_name = ifelse(grepl("TUF TUF", boatName, ignore.case = TRUE),
                        paste0(">>> ", boatName, " <<<"),
                        boatName),
        boat_name = substr(boat_name, 1, 38)
    ) %>%
    head(25) %>%
    select(rank, boat_name, category, speed, dtf, dtl)

gt(df_display) %>%
    cols_label(rank = "Rank", boat_name = "Boat", category = "Class",
               speed = "Speed", dtf = "DTF", dtl = "DTL") %>%
    fmt_number(columns = c(speed, dtf, dtl), decimals = 1) %>%
    tab_options(table.font.size = "small")
```

# Time Trend Analysis

## Since Last Computation

```{r since-last, echo=FALSE, results='asis'}
if (!is.null(last_comp_time) && !is.null(data_since_last) && !is.null(data_since_last$data)) {
    cat(paste0("**Comparison period:** ", data_since_last$hours_ago, " hours ago\n\n"))
    
    last_data <- data_since_last$data
    merged <- merge(current_data, last_data, by = "boat", suffixes = c("_now", "_prev"))
    
    merged$rank_change <- merged$rank_prev - merged$rank_now
    merged$dtf_change <- merged$dtf_now - merged$dtf_prev
    
    changes <- merged[order(merged$rank_change), ]
    top_gainers <- head(changes[changes$rank_change > 0, ], 10)
    top_losers <- head(changes[changes$rank_change < 0, ], 10)
    
    cat("### Biggest Gainers\n\n")
    if (nrow(top_gainers) > 0) {
        gainers_display <- merge(top_gainers, binfo_df, by = "boat") %>%
            select(boatName, rank_now, rank_prev, rank_change, dtf_change) %>%
            rename(Boat = boatName, Current = rank_now, Previous = rank_prev, Change = rank_change, DTF_Diff = dtf_change)
        print(gainers_display)
    } else {
        cat("No rank changes detected.\n")
    }
    
    cat("\n### Biggest Losers\n\n")
    if (nrow(top_losers) > 0) {
        losers_display <- merge(top_losers, binfo_df, by = "boat") %>%
            select(boatName, rank_now, rank_prev, rank_change, dtf_change) %>%
            rename(Boat = boatName, Current = rank_now, Previous = rank_prev, Change = rank_change, DTF_Diff = dtf_change)
        print(losers_display)
    } else {
        cat("No rank changes detected.\n")
    }
} else {
    cat("No previous computation data available for comparison.\n")
}
```

## Last 4 Hours

```{r last-4h, echo=FALSE, results='asis'}
if (!is.null(data_4h) && !is.null(data_4h$data)) {
    cat(paste0("**Reference time:** ", round(data_4h$hours_ago, 1), " hours ago\n\n"))
    
    ref_data <- data_4h$data
    merged_4h <- merge(current_data, ref_data, by = "boat", suffixes = c("_now", "_4h"))
    
    merged_4h$rank_change_4h <- merged_4h$rank_4h - merged_4h$rank_now
    merged_4h$dtf_change_4h <- merged_4h$dtf_now - merged_4h$dtf_4h
    merged_4h$dist_4h <- merged_4h$dtf_4h - merged_4h$dtf_now
    
    changes_4h <- merged_4h[order(merged_4h$rank_change_4h), ]
    top_gainers_4h <- head(changes_4h[changes_4h$rank_change_4h > 0, ], 10)
    
    cat("### Top Gainers (Last 4h)\n\n")
    if (nrow(top_gainers_4h) > 0) {
        gainers_4h_display <- merge(top_gainers_4h, binfo_df, by = "boat") %>%
            select(boatName, rank_now, rank_4h, rank_change_4h, dist_4h) %>%
            rename(Boat = boatName, Current = rank_now, `4h_Ago` = rank_4h, Gain = rank_change_4h, Distance = dist_4h)
        print(gainers_4h_display)
    } else {
        cat("No rank changes in the last 4 hours.\n")
    }
    
    cat("\n### Distance Covered (Last 4h)\n\n")
    top_dist_4h <- merged_4h[order(-merged_4h$dist_4h), ] %>% head(10)
    if (nrow(top_dist_4h) > 0) {
        dist_4h_display <- merge(top_dist_4h, binfo_df, by = "boat") %>%
            select(boatName, rank_now, dist_4h, speed_now) %>%
            rename(Boat = boatName, Rank = rank_now, Distance_4h = dist_4h, Speed = speed_now)
        print(dist_4h_display)
    }
} else {
    cat("No data available for 4-hour comparison.\n")
}
```

## Last 24 Hours

```{r last-24h, echo=FALSE, results='asis'}
if (!is.null(data_24h) && !is.null(data_24h$data)) {
    cat(paste0("**Reference time:** ", round(data_24h$hours_ago, 1), " hours ago\n\n"))
    
    ref_data_24h <- data_24h$data
    merged_24h <- merge(current_data, ref_data_24h, by = "boat", suffixes = c("_now", "_24h"))
    
    merged_24h$rank_change_24h <- merged_24h$rank_24h - merged_24h$rank_now
    merged_24h$dtf_change_24h <- merged_24h$dtf_now - merged_24h$dtf_24h
    merged_24h$dist_24h_actual <- merged_24h$dtf_24h - merged_24h$dtf_now
    
    changes_24h <- merged_24h[order(merged_24h$rank_change_24h), ]
    top_gainers_24h <- head(changes_24h[changes_24h$rank_change_24h > 0, ], 10)
    
    cat("### Top Gainers (Last 24h)\n\n")
    if (nrow(top_gainers_24h) > 0) {
        gainers_24h_display <- merge(top_gainers_24h, binfo_df, by = "boat") %>%
            select(boatName, rank_now, rank_24h, rank_change_24h, dist_24h_actual) %>%
            rename(Boat = boatName, Current = rank_now, `24h_Ago` = rank_24h, Gain = rank_change_24h, Distance = dist_24h_actual)
        print(gainers_24h_display)
    } else {
        cat("No rank changes in the last 24 hours.\n")
    }
    
    cat("\n### Distance Covered (Last 24h)\n\n")
    top_dist_24h <- merged_24h[order(-merged_24h$dist_24h_actual), ] %>% head(10)
    if (nrow(top_dist_24h) > 0) {
        dist_24h_display <- merge(top_dist_24h, binfo_df, by = "boat") %>%
            select(boatName, rank_now, dist_24h_actual, speed_now) %>%
            rename(Boat = boatName, Rank = rank_now, Distance_24h = dist_24h_actual, Speed = speed_now)
        print(dist_24h_display)
    }
} else {
    cat("No data available for 24-hour comparison.\n")
}
```

# TUF TUF Analysis

```{r tuf-info, echo=FALSE, results='asis'}
tuf_row <- subset(results_df, grepl("TUF TUF", boatName, ignore.case = TRUE))

if(nrow(tuf_row) > 0 && "boatName" %in% names(tuf_row)) {
    cat("\n### Info\n\n")
    cat(paste0("- **Name:** ", tuf_row$boatName, "\n"))
    cat(paste0("- **Class:** ", tuf_row$category, "\n"))
    cat(paste0("- **Boat Type:** ", tuf_row$boatClass, "\n"))
    cat(paste0("- **Skipper:** ", tuf_row$skipperNames, "\n\n"))
    cat("\n### Performance\n\n")
    cat(paste0("- **Rank:** ", tuf_row$rank, " / ", nrow(results_df), "\n"))
    cat(paste0("- **Speed:** ", tuf_row$speed, " kts\n"))
    cat(paste0("- **VMG:** ", tuf_row$vmg, " kts\n"))
    cat(paste0("- **DTF:** ", tuf_row$dtf, " nm\n"))
    cat(paste0("- **DTL:** ", tuf_row$dtl, " nm\n"))
    
    if (!is.null(data_4h) && !is.null(data_4h$data)) {
        tuf_4h <- subset(data_4h$data, boat == tuf_row$boat)
        if (nrow(tuf_4h) > 0) {
            cat("\n### 4h Trend\n\n")
            cat(paste0("- **Rank change:** ", tuf_4h$rank - tuf_row$rank, "\n"))
            cat(paste0("- **DTF change:** ", round(tuf_row$dtf - tuf_4h$dtf, 1), " nm\n"))
        }
    }
    
    if (!is.null(data_24h) && !is.null(data_24h$data)) {
        tuf_24h <- subset(data_24h$data, boat == tuf_row$boat)
        if (nrow(tuf_24h) > 0) {
            cat("\n### 24h Trend\n\n")
            cat(paste0("- **Rank change:** ", tuf_24h$rank - tuf_row$rank, "\n"))
            cat(paste0("- **DTF change:** ", round(tuf_row$dtf - tuf_24h$dtf, 1), " nm\n"))
        }
    }
}
```

# Figaro Leaderboard

```{r table2, echo=FALSE}
figaro_boats <- subset(results_df, grepl("Figaro", boatClass, ignore.case = TRUE))

figaro_display <- figaro_boats[order(figaro_boats$rank), ] %>%
    mutate(
        boat_name = ifelse(grepl("TUF TUF", boatName, ignore.case = TRUE),
                        paste0(">>> ", boatName, " <<<"),
                        boatName),
        boat_name = substr(boat_name, 1, 38)
    ) %>%
    head(15) %>%
    select(rank, boat_name, speed, vmg, dtl)

gt(figaro_display) %>%
    cols_label(rank = "Rank", boat_name = "Boat", speed = "Speed", vmg = "VMG", dtl = "DTL") %>%
    fmt_number(columns = c(speed, vmg, dtl), decimals = 1) %>%
    tab_options(table.font.size = "small")
```

---

*Report generated automatically by SailingRaceTrackers*
*Computation times saved to: `r COMPUTATION_TIMES_FILE`*